---
title: "STATS NZ Agricultural Census 2017"
author: "Emma Wang"
date: '2022-07-15'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("readxl", "tidyverse", "ggplot2", "naniar", "sf", "spData", 
               "gganimate", "transformr")
```

# Data Importation and Wrangling
Data Source: Stats NZ Agricultural production statistics: June 2017 (final)  
https://www.stats.govt.nz/information-releases/agricultural-production-statistics-june-2017-final

```{r  dataImportation}
# Define the URL 
URL = "https://www.stats.govt.nz/assets/Uploads/Agricultural-production-statistics/Agricultural-production-statistics-June-2017-final/Download-data/agricultural-production-statistics-jun17-final-tables-v2.xlsx"

# Download the xlsx
download.file(URL, destfile="agricultural-production-statistics-jun17-final-tables-v2.xlsx", quiet = TRUE, mode = "wb")

# Import Data
data_raw <- read_xlsx("agricultural-production-statistics-jun17-final-tables-v2.xlsx", 
                      sheet = 2, skip = 9, col_names = FALSE)

# Remove the unwanted rows and columns
## The ones with complete NAs and Total Numbers
data <- data_raw %>%
  select(seq(1,19, by = 2)) %>%
  filter(complete.cases(.) & !grepl("^Total", `...1`))

# Change column names
name <- expand.grid(x = c('2012_count', '2017_count', '%change'),
            y = c('sheep', 'dairy cattle', 'beef cattle')) %>%
    mutate(nn = paste(x, y, sep = '_')) %>%
    pull(nn)
col_names <- c("Region", name)

names(data) <- col_names


# Deal with character input (S, C or -). Treat them as NA
data <- as.tibble(apply(data, 2, function(x) 
  str_replace_all(x, '^([CS-])$', NA_character_)))

# Change data types in each column
data[all_of(name)] <- sapply(data[name], as.numeric)


```

Now we separate the data into 2 different data sets, because typically percentage changes and counts are not plotted on the same scales.  
```{r dataSep}
data_change <- data %>% select(Region, contains("%change"))
data_count <- data %>% select(Region, contains("Count"))
```

## Convert to long formats
Turn `data_change` into a long format.  
```{r LongFormat1}
data_change <- data_change %>% 
  pivot_longer(!Region, values_to = "Changes", names_to = "Livestock") %>%
  mutate(Livestock = str_replace(Livestock, "%change_", ""),
         Changes = round(Changes, 1)) %>% # Rounding 
  arrange(Livestock)
  
head(data_change)
```

Turn `data_count` into a long format.  
```{r LongFormat2}
data_count <- data_count %>% 
  pivot_longer(!Region, values_to = "Count (thousands)", names_to = "Names") %>%
  separate(2, into = c("Year", "Livestock"), sep = "_count_")
  
head(data_count)
```


# Visual Exploration
## Maps
`nz` is a simple feature data set in `spData` package, which contains the outlines of the 16 regions in North Island and South Island. 
```{r Mergesf}
countsf <- merge(spData::nz, data_count, by.x = "Name", by.y = "Region")
```



```{r}
mapnz <- 
  countsf %>%
  ggplot() + 
  geom_sf() +
  geom_sf_label(data = filter(nz, 
                              !Name %in% c("Tasman", "Taranaki", 
                                           "Bay of Plenty", "Waikato", 
                                           "Manawatu-Wanganui", "Marlborough")),
                mapping = aes(label = Name),
                label.padding = unit(1, "mm"), size = 2) +
  geom_sf_label(data = filter(nz, Name %in% c("Taranaki", "Bay of Plenty")),
                mapping = aes(label = Name),
                label.padding = unit(1, "mm"), size = 2, 
                nudge_y = 20000) +
  geom_sf_label(data = filter(nz, Name %in% c("Waikato", "Tasman")),
                mapping = aes(label = Name),
                label.padding = unit(1, "mm"), size = 2, 
                nudge_x = -50000) +
  geom_sf_label(data = filter(nz, Name %in% c("Waikato")),
                mapping = aes(label = Name),
                label.padding = unit(1, "mm"), size = 2, 
                nudge_x = -50000) +
  geom_sf_label(data = filter(nz, Name == "Manawatu-Wanganui"),
                mapping = aes(label = Name),
                label.padding = unit(1, "mm"), size = 2, 
                nudge_y = -50000) +
  geom_sf_label(data = filter(nz, Name %in% c("Marlborough")),
                mapping = aes(label = Name),
                label.padding = unit(1, "mm"), size = 2, 
                nudge_y = -25000) +
  theme_bw() +
  labs(x = "Latitude", y = "Longitude")

ggsave("Figures/mapnz.jpg", mapnz)
```


* Map of Chatham Islands  
The outlines of Chatham Islands are not included in the `spData` package.  
The map data `nz-chatham-island-island-polygons-topo-150k.csv` were downloaded from Land Information New Zealand:  
https://data.linz.govt.nz/layer/50086-nz-chatham-island-island-polygons-topo-150k/metadata/

```{r}
mapc <- st_read("nz-chatham-island-island-polygons-topo-150k.csv")

mapci <- 
  mapc %>%
  ggplot() +
  geom_sf() +
  geom_label(aes(x = 3513000, y = 5140000, label = "Chatham Islands")) +
  theme(panel.grid.major = element_blank(),
        panel.background = element_rect(fill = "white", colour = "black"), 
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())

# ggsave("Figures/mapci.jpg", mapci)

mapci
```

Below are 2 function to plot the choropleth maps for the 2 main islands and Chatham Islands.  

```{r functions}
# 2 functions to create the choropleth maps
## Main islands
plot_map <- function(live, year = 2017, data = countsf){
  out <- data %>%
    filter(Livestock == live & Year == year) %>%
    ggplot() + 
    geom_sf(aes(fill = `Count (thousands)`)) +
    ggtitle(paste("Total counts of", live, "in", year, "by regions")) +
    theme_bw() +
    labs(x = "Latitude", y = "Longitude") +
    scale_fill_distiller(palette = "Greens",
                         limits = c(min(data %>% 
                                          filter(Year == year & 
                                                   Livestock == live) %>%
                                          pull(`Count (thousands)`), 
                                        na.rm = T),
                                    max(data %>% 
                                            filter(Year == year & 
                                                   Livestock == live) %>%
                                          pull(`Count (thousands)`), 
                                        na.rm = T)))
  return(out)
}

## Chatham Islands
plot_chatham <- function(live, year = 2017, data = mapc, main_data = data_count) {
  # Count of sheep of Chatham Islands
  countc <- data_count %>% 
    filter(Year == year & Livestock == live & 
             Region == "Chatham Islands") %>%
  pull(`Count (thousands)`)

  # Plot the map of Chatham Island
  out <- data %>%
    ggplot() +
    geom_sf(aes(fill = countc)) +
    theme(panel.grid.major = element_blank(),
          panel.background = element_rect(fill = "white", colour = "black"), 
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank()) +
    scale_fill_distiller(name = "Count (thousands)", 
                         palette = "Greens", 
                         limits = c(min(main_data %>% 
                                          filter(Year == year & 
                                                   Livestock == live) %>%
                                          select(`Count (thousands)`), 
                                        na.rm = T),
                                    max(main_data %>% 
                                          filter(Year == year & 
                                                   Livestock == live) %>%
                                          select(`Count (thousands)`), 
                                        na.rm = T)))
  return(out)
}
```


### Sheep
```{r sheepMap}
# Main islands
count_sheep_2017 <- plot_map("sheep")
count_sheep_2017
# Save the map
# ggsave("Figures/count_sheep_2017.jpg", count_sheep_2017)


# Chatham Islands
count_sheep_2017_c <- plot_chatham("sheep")
count_sheep_2017_c

# Save the map
# ggsave("Figures/count_sheep_2017_c.jpg", count_sheep_2017_c)
```


### beef cattle
```{r beefMap}
# Main islands
count_beef_2017 <- plot_map("beef cattle")
count_beef_2017
# Save the map
# ggsave("Figures/count_beef_2017.jpg", count_beef_2017)


# Chatham Islands
count_beef_2017_c <- plot_chatham("beef cattle")
count_beef_2017_c

# Save the map
# ggsave("Figures/count_beef_2017_c.jpg", count_beef_2017_c)
```

### dairy cattle
```{r dairyMap}
# Main islands
count_dairy_2017 <- plot_map("dairy cattle")
count_dairy_2017
# Save the map
# ggsave("Figures/count_dairy_2017.jpg", count_dairy_2017)

# The data for Chatham Islands are unavailable


```



## Changes in counts
### Sheep
```{r sheepCountsNorth}
# A function to plot the animation
plot_anim <- function(live, island, data = data_count, 
                      add_expand = 500){
  north <- c("Northland", "Auckland", "Waikato", "Bay of Plenty", 
           "Gisborne", "Taranaki", "Hawke's Bay",
           "Manawatu-Wanganui", "Wellington")
  south <- c("Nelson", "Tasman", "Marlborough", "West Coast",
             "Canterbury", "Otago", "Southland", "Chatham Islands")
  if(island == "North") {regions <- north
  }else{
    regions <- south
  }
  
  # An animation of livestock counts, North Island
out <- 
  data_count %>% 
  filter(Livestock == live & Region %in% regions) %>%
  ggplot(aes(x = Region, y = `Count (thousands)`)) +
  geom_col(fill = "#778C3C") +
  scale_y_continuous(expand = expansion(add = c(0, add_expand))) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(colour = "grey"),
        panel.background = element_rect(colour = "black", fill = "white"))+
  transition_states(Year, wrap = FALSE) +
  ggtitle(paste("{closest_state}", live, "Counts on the", island, "Island")) +
  enter_fade() +
  exit_fade()

  return(out)
}

# Draw the animation
anim_sheep_north <- plot_anim("sheep", "North")

# Save
# anim_save("Figures/sheep_north.gif", animate(anim_sheep_north, fps = 8, nframes = 64, height = 480, width = 600))
```

```{r sheepCountsSouth}
# Sheep counts for the South island
anim_sheep_south <- plot_anim("sheep", "South")


# anim_save("Figures/sheep_south.gif", animate(anim_sheep_south, fps = 8, nframes = 64, height = 480, width = 600))
```

### Beef Cattle
```{r beefCountsNorth}
# Beef cattle counts on the North island
anim_beef_north <- plot_anim("beef cattle", "North", add_expand = 50)


# anim_save("Figures/beef_north.gif", animate(anim_beef_north, fps = 8, nframes = 64, height = 480, width = 600))
```

```{r beefCountsNorth}
# Beef cattle counts on the North island
anim_beef_south <- plot_anim("beef cattle", "South", add_expand = 50)


# anim_save("Figures/beef_south.gif", animate(anim_beef_south, fps = 8, nframes = 64, height = 480, width = 600))
```




### Dairy Cattle
```{r dairyCountsNorth}
# Beef cattle counts on the North island
anim_dairy_north <- plot_anim("dairy cattle", "North", add_expand = 50)


# anim_save("Figures/dairy_north.gif", animate(anim_dairy_north, fps = 8, nframes = 64, height = 480, width = 600))
```

```{r beefCountsNorth}
# Beef cattle counts on the North island
anim_dairy_south <- plot_anim("dairy cattle", "South", add_expand = 50)


# anim_save("Figures/dairy_south.gif", animate(anim_dairy_south, fps = 8, nframes = 64, height = 480, width = 600))
```